---
title: 'The CLT and the exponential distribution'
author: "JoeriW"
date: "Wednesday, March 11, 2015"
output: html_document
---

###overview
We'll investigate the distribution of 40 variables drawn from an exponential distribution (with a lambda of 0.2)
by comparing the sample mean to the theoretical mean, show how variable the sample is (via variance) and compare it to 
theoretical variance and demonstrate the obtained distribution is approximately normal

###simulation

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. Lambda will be set
**0.2** for all the simulations. The number of simulations is **1000**

set the parameters:
```{r}
nosim = 1000
lambda = 0.2
noexp = 40
```

create a data frame containing the simulated data:
```{r}
dat <- data.frame(
        x = c(apply(matrix(rexp(nosim*noexp,lambda),nosim),1,mean))
        )
```

####Sample Mean versus Theoretical Mean:

calculate the sample mean and the theoretical mean and store them in a data.frame (in order to plot the legend in ggplot)
```{r}
sample.mean <- mean(dat$x)
theo.mean <- 1/lambda

vlines.df <-data.frame(
        mean = c("theoretical","sample"),
        value=c(theo.mean,sample.mean))
```

plot the obtained distribution and add the theoretical and sample mean on the graph:
```{r}
library(ggplot2)
g <- ggplot(data = dat, aes(x = x))
g <- g + geom_density(fill="gold", alpha=0.3, color="black")
g <- g + geom_vline(data = vlines.df
                    ,aes(xintercept = value
                    ,colour = mean
                    ,linetype = mean)
                    ,show_guide = T
                    ,size = 1.1)
g <- g + scale_linetype_manual(values=c("solid", "dotted"))
g <- g + scale_colour_manual(values = c("darkorange1","firebrick1"))
g <- g + labs(title="Distribution")
g
```

When lambda is **`r lambda`**, the theoretical mean of the exponential equals **`r theo.mean`** (= 1/lambda),
the sample mean equals **`r sample.mean`**.
